VOI-aware Monte Carlo Sampling in Trees

### 2011-12-09 paper/vago/paper.tex
paper draft done, all parts in place

Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{description}
\item[Hoeffding inequality \rm{\cite{Hoeffding.ineq}}:] 
\begin{equation}
\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
\label{eqn:conc-hoeffding}
\end{equation}
\item[Empirical Bernstein inequality
  \rm{\cite{MaurerPontil.benrstein}} \textrm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
\label{eqn:conc-empbernstein}
\end{eqnarray}
where sample variance $\overline\sigma_n^2$ is
\begin{equation}
\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
\label{eqn:sample-variance}
\end{equation}
\end{description}
Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.

@@

\textit{TODO}

\appendix

>>>>

Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{description}
\item[Hoeffding inequality \rm{\cite{Hoeffding.ineq}}:] 
\begin{equation}
\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
\label{eqn:conc-hoeffding}
\end{equation}
\item[Empirical Bernstein inequality \rm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
\label{eqn:conc-empbernstein}
\end{eqnarray}
where sample variance $\overline\sigma_n^2$ is
\begin{equation}
\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
\label{eqn:sample-variance}
\end{equation}
\end{description}
Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.

@@

This work suggested Monte Carlo sampling policies in which sample
selection and termination are based on upper bounds on the value of
information. A simplified model of accounting for future value of
information of a sample, based on early termination and sample
redistribution, was proposed for Monte Carlo sampling in
trees. Empirical evaluation showed that the policies outperform
heuristic algorithms for simple regret minimization in multi-armed
bandits, as well as for tree search.

Monte Carlo tree search still remains a largely unexplored field of
application of VOI-aware algorithms. More elaborated VOI estimates,
taking into consideration re-use of samples in future search states
should be considered. The policies introduced in the paper differ from
the UCT algorithm only at the first step, where the VOI-aware
decisions are made. Consistent application of principles of rational
metareasoning at all steps of a rollout may further improve the
sampling policies.
